### 2013-10-24 paper/paper.tex
creating folder for the paper

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}

\oddsidemargin 0in
\evensidemargin 0in
\textwidth= 6.5in
\topmargin -0.50in
\textheight 9in

\usepackage{latexsym,amssymb,amsmath,color}
\newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}
                     %

\begin{document}
\begin{center}{\Large\bf Tsunami Uncertainty Quantification}\vskip 5pt

\bf Preliminary results
\end{center}
\vskip 15pt

\input{intro}

\input{tsunami}

\clearpage
%\input{inference}

\input{uqpce}
\input{results}

\clearpage

\clearpage
\bibliographystyle{plainnat}
\bibliography{biblio}
\end{document}