### 2013-10-24 paper/results.tex
adding tex files

\section{Results}
\label{sec:results}
The results of the forward and inverse problems are presented and 
discussed in three sections. First, we perform an error analysis 
of the PC expansion construction, namely  to establish the 
validity of the corresponding surrogate.  Second, we perform a 
sensitivity analysis and rank the impact of the different Manning's 
roughness coefficient on the predicted surface elevation. Finally, 
posterior distributions for the input parameters are determined 
and analyzed in light of the available gauge data.
\input{pcstudy}
\input{analysis}
\input{mcmc}