### 2013-10-30 paper/cover.tex
adding inference section

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\begin{center}
\begin{Large}
{\bf Polynomial Chaos for the Estimation of Manning's Friction}\\

\end{Large}
\bigskip
\bigskip
Ihab Sraj$^1$, Kyle Mandli$^2$ \& Ibrahim Hoteit$^1$\\
\bigskip
$^1$Department of Physical Sciences and Engineering,
King Abdullah University for Science and Technology, Thuwal, Saudi Arabia \\
$^2$Department of Engineering, Austin, TX 
\end{center}

\vspace{5.5cm}

\begin{tabbing}
Corresponding Author: \hspace{5mm} \=  \\
       \> Department of Physical Sciences and Engineering \\
       \> King Abdullah University for Science and Technology \\
       \> Thuwal, Saudi Arabia\\
       \> \\
Phone: \>  \\
Fax:   \>  \\
Email: \>    \\
\\
Submitted to: \> {\it TBD} \\
\> \today \\
\\
Revised: \>

\end{tabbing}