### 2013-11-07 paper/inference.tex
major changes to formulation part

\subsection{Bayesian Inference}
 \label{sec:inference}

Bayesian inference is a statistical approach to inverse problems
that has recently gained much interest in different applications including
ocean~\citep{Alexanderian2011a,Zedler2012,sraj:2013a}
climate~\citep{OlsonEtAl2012} and geophysical~\citep{Malinverno2002} modeling.
The key idea in this approach is to express all forms of uncertainty
in terms of probability. The first step in a Bayesian inference 
is to formulate the forward problem (response surface/model) using 
a suitable likelihood function and a product of conditional probability densities. 
We then formulate the prior distribution over the unknown parameters of the model
using our best belief. Lastly, given some observation data, Bayes rule 
is used to obtain a posterior distribution for these unknown
\citep{sivia}. We review this approach briefly below.

Let $\vec{d}=(d_1,...,d_m)^T$ be a vector of observation data and $\vec{\theta}=(\theta_1,...,\theta_n)^T$ be a vector of model parameters or inputs. We now consider a forward model $\vec G$ that predicts the data function of 
the parameters such that:
\begin{equation}
\vec d \approx \vec{G}( \vec \theta)
\end{equation}

Applying the Bayes' rule yields:
\begin{equation}
 p(\vec{\theta}| \vec d) \propto 
 p(\vec d | \vec{\theta}) \ p(\vec{\theta})  
\label{eq:bayes}
\end{equation}
where $p(\vec{\theta})$ is the prior of $\vec{\theta}$, representing the \emph{a priori} knowledge
about the parameters; 
$p(\vec d| \vec{\theta})$ is the likelihood function representing
the probability of obtaining the data given the set of parameters $\vec{\theta}$;
and finally $p(\vec{\theta}| \vec d)$ is the posterior,
representing the probability that $\vec{\theta}$ is true given the data $\vec d $.

To formulate the likelihood function, we let $\vec \epsilon = \vec d - \vec{G}$
represents the discrepancy between the model and observations.
Here, the components of $\vec \epsilon $ are i.i.d. random variables with density $p_{\epsilon}$.
In this case the likelihood function can be written as:
\begin{equation} 
p(\vec d |  \vec{\theta}) 
= 
\prod_{i=1}^m  
p_\epsilon (d_i - G_i(\vec \theta))  	
\label{eq:likelihood}
\end{equation}

In our application, we assume that the errors $\epsilon_i$ are independent
and normally distributed with mean
zero and variance $\sigma^2$, i.e. 
$\epsilon_i \sim N(0,\sigma^2)$. 
While in general $\sigma^2$ depends on the observations, in cases where the error
amplitude is generally small and does not change throughout space and time, one may use
a simplified model and assume single $\sigma^2$ value.
Consequently, Equation~\ref{eq:likelihood} becomes:
\begin{equation} 
p(\vec d |  \vec{\theta}) 
= 
\frac{1}{\sqrt{2 \pi \sigma^2}}\prod_{i=1}^m   
\exp \left\lbrace \frac{-(d_i - G_i(\vec \theta))^2}{2 \sigma^2} \right\rbrace 	
\label{eq:likelihood2}
\end{equation}
The joint posterior from Bayes' rule can be given as:
\begin{equation} 
p(\vec{\theta}| \vec d)
\propto
\frac{1}{\sqrt{2 \pi \sigma^2}}   \prod_{i=1}^m  
\exp \left\lbrace \frac{-(d_i - G_i(\vec \theta))^2}{2 \sigma^2} \right\rbrace  
\prod_{i=1}^n p(\theta_i)
\end{equation}
The variance $\sigma^2$ is unknown \emph{a priori}; thus we treat it as a hyper-parameter.
In other words, $\sigma^2$ may become an additional object for Bayesian inference and may themselves be endowed with priors and estimated from data. In this case the joint posterior is finally written as follows:
\begin{equation} 
p(\vec{\theta},\sigma^2 | \vec d)
\propto
\frac{1}{\sqrt{2 \pi \sigma^2}}   \prod_{i=1}^m  
\exp \left\lbrace \frac{-(d_i - G_i(\vec \theta))^2}{2 \sigma^2} \right\rbrace
\ p(\sigma^2) \prod_{i=1}^n p(\theta_i)
\label{eq:post}
\end{equation}