### 2013-12-11 paper/results.tex

\section{Results}
\label{sec:results}
We begin the results section by performing
an error and convergence analysis of the
constructed surrogate using the PC expansions
to establish its validity in Section~\ref{sec:analysis}.
Then in Section~\ref{sec:forward}, we present
statistical analysis to quantify the uncertainty in the predicted surface elevation
as well as a sensitivity analysis to rank the impact of the different Manning's 
roughness coefficients on this uncertainty. 
Finally, in Section~\ref{sec:inverse}, we present the results of the inverse
problem where we determined the posterior distributions of the input parameters 
and analyzed them in light of the available gauge data.
\input{pcstudy}
\input{analysis}
\input{mcmc}