### 2013-12-16 paper/pcstudy.tex

\subsection{Error and convergence study}
\label{sec:analysis}

Figure~\ref{fig:rlzs} shows the evolution of the
water surface elevation predicted at the four different gauges 
using \geoclaw for the 125 different realizations 
required to compute the PC expansions. We notice that the 
variability in water surface elevation is insignificant for the first 
hour at all gauges as the plots of the different realizations superimpose.
Later, the variability starts to increase with time at the different gauges 
as indicated from the thickness of the bands formed by the plots of the realizations
and is significant at $t\sim$ 5400~s, 6000~s, 3600~s and 7200~s
for gauges 21401, 21413, 21418, 21419 respectively.
This is consistent with the distance from the gauge to the source of the earthquake
where gauge 21418 is the closest and gauge 21419 is the farthest.
The uncertainty in the water surface  elevation continues till the end of the simulations
at all gauges.

In order to check the consistency of the PC approximation, we compare
water surface elevation from the realizations 
with those obtained from the PC surrogate. The different curves
show an excellent agreement (not shown) for all times and gauges locations.
To quantify this agreement, we define
an error metric that measures the relative normalized root mean-square error between the left hand side function 
in Equation~(\ref{eq:stochseries}) and its PC representation at the sampling points:

\begin{equation} 
   E = \frac{\displaystyle
         \left(\sum_{\xxi \in \NISP} \left|U(\xxi) - \sum_{k = 0}^{P}
U_k\Psi_k(\xxi)\right|^2
         \right)^{1/2}}
        {\displaystyle
          \left(\sum_{\xxi \in \NISP} \left|U(\xxi)\right|^2\right)^{1/2} 
          },
\label{eq:error}
\end{equation}
where $\NISP$ is the 125-member ensemble obtained using the PC algorithm. 
This error metric calculated at the different gauge locations is shown in Figure~\ref{fig:error};
the largest relative normalized error for 
water surface elevation is about 0.1\%.

%Contour maps of the relative normalized error for the entire simulation region are 
%shown in Figure~\ref{fig:error2D} for various depths and dates to confirm the error trends of the
%analysis box.  The error is largest after the wind
%intensifies (bottom row) for all depths. For either day, the maximum
%error is located at 50~m which coincides with the depth of the
%original mixed layer measured by the AXBT. The maximum magnitude
%recorded is about 1\% and occurs on Sep~18. The
%elevated error region is located to the right of the storm and
%extends from the surface down to 50~m, after which the impact of the
%input uncertainty decreases substantially.  For the purpose of our
%current study, and given that the majority of AXBT data are at
%depth, the surrogate's errors are considered acceptable and small.

A final check consists of verifying whether the probability density
functions (pdfs) of water surface elevation at the different gauge locations
converges with increased order of the PC representation.  Sample
water surface elevation pdfs are shown in Figure~\ref{fig:pdfs2}
and Figure~\ref{fig:pdfs3}
The different curves correspond to increased order of PC (from 1 to 5).
The plots indicate that the double peaked distributions are
well-resolved with order 3 at t = 7200 s as shown in Figure~\ref{fig:pdfs2}.
However, at t = 10800 s further refinement is needed as the pdfs are
sensitive to the refinement up to order 4 but then becomes weakly 
insensitive with additional refinement  (order 5) as indicated in Figure~\ref{fig:pdfs3}. 
We, therefore, used PC order 5 in all computations.
This test and the various error metrics presented above provide confidence that the PC expansion is a faithful 
model surrogate that can be used in both the forward and inverse problems..