### 2013-12-23 paper/paper.tex

%&pdflatex
%
%
%
%

% \documentclass[12pt]{article}
\documentclass[preprint,12pt]{elsarticle}

\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}
\usepackage{xspace} % Incorporates correct optional space after commands
\usepackage{hyperref}

% Add line numbers
\usepackage[mathlines]{lineno} % Control line numbers

% Multipart figures
\usepackage{subcaption}  % "Newest" subfigure like package

% \oddsidemargin 0in
% \evensidemargin 0in
% \textwidth= 6.5in
% \topmargin -0.50in
% \textheight 9in

\usepackage{ifpdf} % For declaring rules for graphics below

\usepackage{latexsym,amssymb,amsmath,color}
\newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\alert}[1]{\textbf{\textcolor{red}{{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}

\newcommand{\pd}{\pard}
\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}
\newcommand{\dx}{\ensuremath{\Delta x}}   % Delta x
\newcommand{\dy}{\ensuremath{\Delta y}}   % Delta y
\newcommand{\dt}{\ensuremath{\Delta t}}   % Delta t

\journal{Ocean Modelling}

\begin{document}

\ifpdf
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .tif}
\else
\DeclareGraphicsExtensions{.png, .jpg, .tif, .eps}
\fi

\begin{frontmatter}

\title{Estimation of Manning's Friction Coefficient using Polynomial Chaos and Bayesian inference}
\date{\today}

\author[kaust]{Ihab Sraj}
\ead{ihab.sraj@kaust.edu.sa}
\ead{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\author[ut]{Kyle T. Mandli}
% \ead{kyle@ices.utexas.edu}
% \ead[http://users.ices.utexas.edu/~kyle]{http://users.ices.utexas.edu/~kyle}
\author[duke]{Omar M. Knio}
\author[ut]{Clint N. Dawson}
\author[kaust]{Ibrahim Hoteit}

\address[kaust]{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[ut]{Institute for Computational Engineering and Science, University of Texas at Austin, 201 E 24th ST. Stop C0200, Austin, TX 78712-1229, USA}
\address[duke]{Department of Mechanical Engineering and Materials Science, Duke University, 144
Hudson Hall, Durham, North Carolina 27708, USA}
\date{\today}

\begin{abstract}

Tsunamis have been lately catastrophic due to the warnings
issued late and to poorly predicted water elevations. Computational 
models of tsunami events are often employed to predict water elevations 
by exploring multiple different scenarios due to the abundance of uncertainty 
in the input data. Accurate estimation of water elevations requires 
accurate estimation of many model parameters. Our objective here is to 
present an efficient method to estimate the Manning's n friction coefficient
that plays an important role in the tsunami governing equations. 
In our work, we perform a sensitivity analysis of the Manning’s n coefficient 
and infer using water surface elevation data obtained during \tohoku tsunami. 
We characterize the Manning’s friction coefficient by three different parameters 
assumed constant in three regions: $n_1$ on-shore, $n_2$ near-shore and $n_3$ deep-water. 
The efficiency of our approach stems from the use of polynomial chaos expansions to build an inexpensive surrogate for the numerical tsunami \geoclaw model that can be used to perform the sensitivity analysis. The surrogate also reduces the computational burden of the Markov Chain Monte-Carlo sampling needed for the Bayesian inference. Our objective is to sharpen three initial estimates of the three uncertain parameters. Our results indicate that Manning’s n friction coefficients have a Maximum-A-Posteriori (MAP) values of $n_2=0.011$ and $n3=0.185$ while for $n_1$ no meaningful MAP value can be determined using the available data.

\end{abstract}

\begin{keyword}
tsunami \sep Manning's n friction coefficient \sep sensitivity analysis \sep polynomial chaos \sep Bayesian inference
\end{keyword}

\end{frontmatter}
\linenumbers

% \input{cover}
% \clearpage

\input{intro}

\input{tsunami}

\input{formulation}

\input{results}
\clearpage

\input{figures}

\clearpage
\bibliographystyle{elsarticle-num}
\bibliography{biblio}
\end{document}