### 2013-12-23 paper/paper.tex
title and abstract revisited

%&pdflatex
%
%
%
%

% \documentclass[12pt]{article}
\documentclass[preprint,12pt]{elsarticle}

\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}
\usepackage{xspace} % Incorporates correct optional space after commands
\usepackage{hyperref}

% Add line numbers
\usepackage[mathlines]{lineno} % Control line numbers

% Multipart figures
\usepackage{subcaption}  % "Newest" subfigure like package

% \oddsidemargin 0in
% \evensidemargin 0in
% \textwidth= 6.5in
% \topmargin -0.50in
% \textheight 9in

\usepackage{ifpdf} % For declaring rules for graphics below

\usepackage{latexsym,amssymb,amsmath,color}
\newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\alert}[1]{\textbf{\textcolor{red}{{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}

\newcommand{\pd}{\pard}
\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}
\newcommand{\dx}{\ensuremath{\Delta x}}   % Delta x
\newcommand{\dy}{\ensuremath{\Delta y}}   % Delta y
\newcommand{\dt}{\ensuremath{\Delta t}}   % Delta t

\journal{Ocean Modelling}

\begin{document}

\ifpdf
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .tif}
\else
\DeclareGraphicsExtensions{.png, .jpg, .tif, .eps}
\fi

\begin{frontmatter}

\title{Estimation of Manning's Friction Coefficient using DART Buoys Data during
\tohoku Tsunami}
\date{\today}

\author[kaust]{Ihab Sraj}
\ead{ihab.sraj@kaust.edu.sa}
\ead{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\author[ut]{Kyle T. Mandli}
% \ead{kyle@ices.utexas.edu}
% \ead[http://users.ices.utexas.edu/~kyle]{http://users.ices.utexas.edu/~kyle}
\author[duke]{Omar M. Knio}
\author[ut]{Clint N. Dawson}
\author[kaust]{Ibrahim Hoteit}

\address[kaust]{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[ut]{Institute for Computational Engineering and Science, University of Texas at Austin, 201 E 24th ST. Stop C0200, Austin, TX 78712-1229, USA}
\address[duke]{Department of Mechanical Engineering and Materials Science, Duke University, 144
Hudson Hall, Durham, North Carolina 27708, USA}
\date{\today}

\begin{abstract}

Tsunamis have been lately catastrophic due to the warnings
issued late and to the inaccurate predictions of water elevations. 
Tsunami computational models are often employed to explore multiple 
different scenarios to predict water elevations due to the abundance 
of uncertainty in the input data. However, accurate estimation of water elevations 
requires accurate estimation of many model parameters,
in particular the Manning's $n$ friction coefficient that plays an important role in the 
tsunami governing equations. Our objective here is to present an efficient method for the 
estimation of Manning's $n$ coefficient characterized by three different parameters 
assumed constant in three regions: $n_1$ on-shore, $n_2$ near-shore and $n_3$ deep-water. 
This method uses Polynomial Chaos (PC) to build an inexpensive surrogate for the 
\geoclaw model and then employs the Bayesian inference that uses the DART buoys data
obtained during \tohoku tsunami to estimate the three uncertain parameters. The efficiency of our approach stems from 
using a surrogate model that reduces the computational burden of the Markov Chain Monte-Carlo 
sampling needed in the Bayesian inference. The PC surrogate is also used to perform a sensitivity analysis. 
Our results indicate that Manning’s $n$ coefficients have a Maximum-A-Posteriori (MAP) values of $n_2=0.011$ and $n_3=0.180$, while for $n_1$ no meaningful MAP value can be determined using the available data.

\end{abstract}

\begin{keyword}
tsunami \sep Manning's n friction coefficient \sep sensitivity analysis \sep polynomial chaos \sep Bayesian inference
\end{keyword}

\end{frontmatter}
\linenumbers

% \input{cover}
% \clearpage

\input{intro}

\input{tsunami}

\input{formulation}

\input{results}
\clearpage

\input{figures}

\clearpage
\bibliographystyle{elsarticle-num}
\bibliography{biblio}
\end{document}