### 2013-12-23 paper/formulation.tex

\section{Formulation}
\label{sec:formu}
In this section, we introduce the methods employed in this work.
First, the Bayesian approach to inverse problems is described in Section~\ref{sec:inference};
its application to our problem of inferring the Manning's $n$ coefficients in Section~\ref{sec:manning}.
Finally, in Section~\ref{sec:uqpce}, we introduce
a stochastic spectral method for forward uncertainty propagation and accelerating the Bayesian inference.

\input{inference}
\input{coefs}

\input{uqpce}