### 2013-12-23 paper/results.tex

\section{Results}
\label{sec:results}
We begin the results section by performing an error and convergence 
analysis of the constructed surrogate using the PC expansions to 
establish its validity in Section~\ref{sec:analysis}.  In Section
~\ref{sec:forward}, we present a statistical analysis to quantify the 
uncertainty in the predicted water surface elevation in addition to a 
sensitivity analysis that enables us to rank the impact of the 
different Manning's $n$ coefficients on the uncertainty. Finally, 
in Section~\ref{sec:inverse}, we present the results of the inverse problem 
where we determine the posterior distributions of the input uncertain
parameters and analyze them in light of the available gauge data.
\input{pcstudy}
\input{analysis}
\input{mcmc}