### 2013-12-23 paper/conc.tex
added outline

\section{Discussion and conclusions}
\label{sec:conc}