### 2014-01-08 paper/formulation.tex

\section{Formulation}
\label{sec:formu}
In this section, we first, present the Bayesian approach to inverse problems 
and its application to our problem of inferring the Manning's $n$ coefficients.
We also introduce
a stochastic Polynomial Chaos-based spectral method for forward uncertainty propagation and accelerating the Bayesian inference.

\input{inference}
\input{coefs}

\input{uqpce}