### 2014-01-08 paper/paper.tex

%&pdflatex
%
%
%
%

% \documentclass[12pt]{article}
\documentclass[preprint,12pt]{elsarticle}

\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}
\usepackage{xspace} % Incorporates correct optional space after commands
\usepackage{hyperref}

% Add line numbers
\usepackage[mathlines]{lineno} % Control line numbers

% Multipart figures
\usepackage{subcaption}  % "Newest" subfigure like package

% \oddsidemargin 0in
% \evensidemargin 0in
% \textwidth= 6.5in
% \topmargin -0.50in
% \textheight 9in

\usepackage{ifpdf} % For declaring rules for graphics below

\usepackage{latexsym,amssymb,amsmath,color}
\newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\alert}[1]{\textbf{\textcolor{red}{{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}

\newcommand{\pd}{\pard}
\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}
\newcommand{\dx}{\ensuremath{\Delta x}}   % Delta x
\newcommand{\dy}{\ensuremath{\Delta y}}   % Delta y
\newcommand{\dt}{\ensuremath{\Delta t}}   % Delta t

\journal{Ocean Modelling}

\begin{document}

\ifpdf
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .tif}
\else
\DeclareGraphicsExtensions{.png, .jpg, .tif, .eps}
\fi

\begin{frontmatter}

\title{Uncertainty Quantification and Inference of Manning's Friction Coefficient using DART Buoys Data during
\tohoku Tsunami}
\date{\today}

\author[kaust]{Ihab Sraj}
\author[ut]{Kyle T. Mandli}
% \ead{kyle@ices.utexas.edu}
% \ead[http://users.ices.utexas.edu/~kyle]{http://users.ices.utexas.edu/~kyle}
\author[duke]{Omar M. Knio}
\author[ut]{Clint N. Dawson}
\author[kaust]{Ibrahim Hoteit}
\ead{ibrahim.hoteit@kaust.edu.sa}
\ead{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[kaust]{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[ut]{Institute for Computational Engineering and Science, University of Texas at Austin, 201 E 24th ST. Stop C0200, Austin, TX 78712-1229, USA}
\address[duke]{Department of Mechanical Engineering and Materials Science, Duke University, 144
Hudson Hall, Durham, North Carolina 27708, USA}
\date{\today}

\begin{abstract}

%Tsunamis have been lately catastrophic due to the warnings
%issued late and to the inaccurate predictions of water elevations. 
Tsunami computational models are employed to explore multiple 
different scenarios and to predict water elevations 
%due to the abundance of uncertainty in the input data. 
However, accurate estimation of water elevations 
requires accurate estimation of many model parameters,
in particular the Manning's $n$ friction coefficient. 
Our objective here is to develop an efficient approach for the 
uncertainty quantification and inference of Manning's $n$ coefficient 
which we characterize here by three different parameters 
as being constant over three different regions: $n_1$ on-shore, $n_2$ near-shore and $n_3$ deep-water. 
We use Polynomial Chaos (PC) to build an inexpensive surrogate for the 
\geoclaw model and then employs Bayesian inference to estimate 
and quantify the uncertainties to those uncertain parameters
using the DART buoys data collected during \tohoku tsunami. 
The surrogate model significantly reduces the computational burden of the Markov Chain Monte-Carlo (MCMC) 
sampling of the Bayesian inference. The PC surrogate is also used to perform a sensitivity analysis. 
Our results indicate that Manning's $n$ coefficients have a Maximum-A-Posteriori (MAP) values of $n_2=0.011$ and $n_3=0.180$, while for $n_1$ no meaningful MAP value can be determined from the available data.

\end{abstract}

\begin{keyword}
tsunami \sep Manning's $n$ friction coefficient \sep sensitivity analysis \sep polynomial chaos \sep Bayesian inference
\end{keyword}

\end{frontmatter}
\linenumbers

% \input{cover}
% \clearpage

\input{intro}

\input{tsunami}

\input{formulation}

\input{results}
\input{conc}

\clearpage

\input{figures}

\clearpage
\bibliographystyle{elsarticle-num}
\bibliography{biblio}
\end{document}