### 2014-01-16 paper/formulation.tex

\section{Formulation}
\label{sec:formu}
In this section, we first, present the Bayesian approach to inverse problems 
and its application to our problem of inferring the Manning's $n$ coefficients.
We also introduce
a Polynomial Chaos method used for forward uncertainty propagation, and 
for accelerating the Bayesian inference based on the resulting surrogate.

\input{inference}
\input{coefs}

\input{uqpce}