### 2014-01-16 paper/results.tex

\section{Results}
\label{sec:results}
We begin by performing an error and convergence 
analysis of the constructed surrogate using the PC expansions to 
establish its validity. We then present a statistical analysis to quantify the 
uncertainty in the predicted water surface elevation, as well as a 
sensitivity analysis that enables us to rank the impact of the 
different Manning's $n$ coefficients on the uncertainty. Finally, 
we present the results of the inverse problem 
where we determine the posterior distributions of the input uncertain
parameters and analyze them in light of the available gauge data.
\input{pcstudy}
\input{analysis}
\input{mcmc}