### 2014-02-03 friction_paper/paper.tex
minor and changes spacing

%&pdflatex
%
%
%
%

% \documentclass[12pt]{article}
\documentclass[review,12pt]{elsarticle}

\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}
\usepackage{xspace} % Incorporates correct optional space after commands
\usepackage{hyperref}

% Add line numbers
\usepackage[mathlines]{lineno} % Control line numbers

% Multipart figures
\usepackage{subcaption}  % "Newest" subfigure like package

% \oddsidemargin 0in
% \evensidemargin 0in
% \textwidth= 6.5in
% \topmargin -0.50in
% \textheight 9in

\usepackage{ifpdf} % For declaring rules for graphics below

\usepackage{latexsym,amssymb,amsmath,color}
% \newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
% \newcommand{\comment}[1]{\textcolor{red}{{#1}}}
% \newcommand{\alert}[1]{\textbf{\textcolor{red}{{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}

\newcommand{\pd}{\pard}
\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}
\newcommand{\dx}{\ensuremath{\Delta x}}   % Delta x
\newcommand{\dy}{\ensuremath{\Delta y}}   % Delta y
\newcommand{\dt}{\ensuremath{\Delta t}}   % Delta t

\journal{Ocean Modelling}

\begin{document}

\ifpdf
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .tif}
\else
\DeclareGraphicsExtensions{.png, .jpg, .tif, .eps}
\fi

\begin{frontmatter}

\title{Uncertainty Quantification and Inference of Manning's Friction Coefficient using DART Buoys Data during
\tohoku Tsunami}
\date{\today}

\author[kaust1]{Ihab Sraj}
\author[ut]{Kyle T. Mandli}
\author[duke,kaust2]{Omar M. Knio}
\author[ut]{Clint N. Dawson}
\author[kaust1]{Ibrahim Hoteit\corref{cor1}}
\ead{ibrahim.hoteit@kaust.edu.sa}
%\ead{Division of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[kaust1]{Division of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}

\address[ut]{Institute for Computational Engineering and Science, University of Texas at Austin, 201 E 24th ST. Stop C0200, Austin, TX 78712-1229, USA}
\address[duke]{Department of Mechanical Engineering and Materials Science, Duke University, 144
Hudson Hall, Durham, North Carolina 27708, USA}
\address[kaust2]{Division of Computer, Electrical and Mathematical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\cortext[cor1]{Corresponding author}
\date{\today}

\begin{abstract}

Tsunami computational models are employed to explore multiple
scenarios and to predict water elevations.  However, accurate estimation of
water elevations  requires accurate estimation of many model parameters, in
particular the Manning's $n$ friction coefficient.  Our objective here is to
develop an efficient approach for the  uncertainty quantification and inference
of the Manning's $n$ coefficient which we characterize here by three different
parameters as being constant over three different regions: $n_1$ on-shore,
$n_2$ near-shore and $n_3$ deep-water.  We use Polynomial Chaos (PC) to build an
inexpensive surrogate for the  \geoclaw model and then employ Bayesian inference
to estimate  and quantify the uncertainties to those uncertain parameters using
the DART buoy data collected during the \tohoku tsunami.  The surrogate model
significantly reduces the computational burden of the Markov Chain Monte-Carlo
(MCMC)  sampling of the Bayesian inference. The PC surrogate is also used to
perform a sensitivity analysis.  Our results indicate that Manning's $n$
coefficients have a Maximum-A-Posteriori (MAP) values of $n_2=0.011$ and
$n_3=0.180$, while for $n_1$ no meaningful MAP value can be determined from the
available data.

\end{abstract}

\begin{keyword}
tsunami \sep Manning's $n$ friction coefficient \sep sensitivity analysis \sep polynomial chaos \sep Bayesian inference
\end{keyword}

\end{frontmatter}
\linenumbers

% \input{cover}
% \clearpage

\input{intro}

\input{tsunami}

\input{formulation}

\input{results}
\input{conc}

\bibliographystyle{elsarticle-num}
\bibliography{biblio}
\clearpage
\input{figures} 
\end{document}