### 2014-04-07 friction_revised/highlights.tex
created folder for revised version of the paper

 \documentclass[12pt]{article}
\usepackage{xspace} % Incorporates correct optional space after commands
\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}

\begin{document}

\begin{itemize}
\item  We present an efficient approach for the inference
of the Manning's $n$ coefficient.
\item  We propose a three-parameter representation of Manning's n coefficient.
\item We use Polynomial Chaos to build an inexpensive surrogate for the \geoclaw model.
\item We use Bayesian inference to infer the parameters using data from \tohoku tsunami.
\item We report \textit{pdfs} and Maximum-A-Posteriori values of the uncertain parameters.
\end{itemize}
\end{document}